Synthesizing Symmetric Lenses

### 2018-05-24 papers/SymmetricLenses/main.tex
add better title
better examples

\title{Simple Symmetric Lenses - Foundations and Synthesis}

@@

% begin introduction
\section{Introduction}
% state of the world
Similar data can come in varieties of formats: bibliography information can be
stored in the BibTeX, EndNote, or ACM reference formats; web service resources
can be provided to clients in legacy or modern formats; calendar events can be
stored in CSV files or ICS files. Oftentimes, we want to synchronize files
across formats: when my collaborator edits her bibliography stored in the ACM
reference format, I want these edits reflected in my EndNote file; when I update
data through a legacy API, I want those changes to be reflected when I retrieve
it through a modern API; when I create a new calendar event in my ICS file, I
want my CSV calendar files to be updated to reflect this change.

@@

A better approach is to synthesize such conversion functions. Recent research
has gone into automatically synthesizing such conversion functions from
input-output examples and formal descriptions of the data
formats~\cite{optician,quotient-optician}. However, this work assumes that all
data fields present in one format are present in the other, but this is not
always true. Oftentimes, fields present in one format are not present in the
other: updated APIs can expose new information, or omit legacy information; the
ACM reference format includes article number information, whereas EndNote
includes page number information; ICS files contain a URL field that Google
Calendar CSV files do not expose. We aim to synthesize transformations where the
data fields are not necessarily in bijective correspondence, where each data
format can contain important information not present in the other format.

@@

In summary, our contributions are:
\begin{enumerate}
\item We develop \emph{simple symmetric lenses} a new formulation for
  bidirectional programs that can project data from both sides~(\S\ref{sec:ssl}),
  and describe a language that expresses such lenses on string
  data~(\S\ref{sec:ssl-string}).
\item We develop a means to traverse stochastic regular expression equivalences,
  while preserving probability distributions and
  languages~(\S\ref{sec:preliminaries}).
\item We develop stochastic DNF regular expressions and DNF lenses. We develop a
  sound algorithm for converting stochastic regular expressions to stochastic
  DNF regular expressions~(\S\ref{sec:dnf-regex-lens}), and a
  information-theoretic cost metric on DNF
  lenses~(\S\ref{sec:symmetric-dnf-lenses}).
\item We extended Boomerang with our simple symmetric lens combinators. We
  updated the Optician synthesis system to synthesize symmetric lenses with our
  new algorithm.
\item We evaluate our synthesis algorithm on Opticians benchmark suite as well
  as new, symmetric benchmarks, and find that we lose very little synthesis
  time, while gaining a large amount of expressivity (\S\ref{sec:evaluation}).
\end{enumerate}
% end introduction

% begin preliminaries
\section{Preliminaries}
\label{sec:preliminaries}
To define the entropy for data recovery, we must first develop a probability
model for our languages. We do this by enhancing regular expressions into
\emph{stochastic regular expressions}. With this, we can jointly
express both a language and a probability distribution over that language.

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}

% end preliminaries

\section{Simple Symmetric Lenses}
\label{sec:ssl}
Classically, asymmetric lenses comprised three functions, \Put{}, \Get{}, and
\Create{}.  We extend this style of lens to symmetric lenses with
\emph{simple symmetric lenses}.  Like the classical formulation of asymmetric
lenses, simple symmetric lenses have no complement.  Simple symmetric
lenses comprise 4 functions, \CreateR, \CreateL, \PutR, and \PutL.
These functions satisfy the following round-tripping laws:
\begin{equation}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x
\end{equation}
\begin{equation}
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y
\end{equation}
\begin{equation}
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{equation}

>>>>

\title{Synthesizing Symmetric Lenses}

@@

% begin introduction
\section{Introduction}
% state of the world
Similar data can come in varieties of formats: scheduled jobs can be stored as
cron jobs for Linux or as Scheduled Tasks for Windows; web service resources can
be provided to clients in legacy or modern formats; calendar events can be
stored in CSV files or ICS files. Oftentimes, we want to synchronize files
across formats: when I edit a task to run more often on my Linux machine, I want
this task to run more often on my Windows machine; when I update data through a
legacy API, I want those changes to be reflected when I retrieve it through a
modern API; when I create a new calendar event in my ICS file, I want my CSV
calendar files to be updated to reflect this change.

@@

A better approach is to synthesize such conversion functions. Recent research
has gone into automatically synthesizing such conversion functions from
input-output examples and formal descriptions of the data
formats~\cite{optician,quotient-optician}. However, this work assumes that all
data fields present in one format are present in the other, but this is not
always true. Oftentimes, fields present in one format are not present in the
other: the commands scheduled to run on Linux machines are typically different
than the commands scheduled to run on Windows machines; updated APIs can expose
new information, or omit legacy information; ICS files contain a URL field that
Google Calendar CSV files do not expose. We aim to synthesize transformations
where the data fields are not necessarily in bijective correspondence, where
each data format can contain important information not present in the other
format.

@@

In summary, our contributions are:
\begin{enumerate}
\item We develop \emph{simple symmetric lenses} a new formulation for
  bidirectional programs that can project data from both sides~(\S\ref{sec:ssl}),
  and describe a language that expresses such lenses on string
  data~(\S\ref{sec:ssl-string}).
\item We develop a means to traverse stochastic regular expression equivalences,
  while preserving probability distributions and
  languages~(\S\ref{sec:stoch-rx}).
\item We develop stochastic DNF regular expressions and DNF lenses. We develop a
  sound algorithm for converting stochastic regular expressions to stochastic
  DNF regular expressions~(\S\ref{sec:dnf-regex-lens}), and a
  information-theoretic cost metric on DNF
  lenses~(\S\ref{sec:symmetric-dnf-lenses}).
\item We extended Boomerang with our simple symmetric lens combinators. We
  updated the Optician synthesis system to synthesize symmetric lenses with our
  new algorithm.
\item We evaluate our synthesis algorithm on Opticians benchmark suite as well
  as new, symmetric benchmarks, and find that we lose very little synthesis
  time, while gaining a large amount of expressivity (\S\ref{sec:evaluation}).
\end{enumerate}
% end introduction

% begin extended example
\section{Extended Example}
We motivate our problem with an extended example.  Consider the previously
described scenario of keeping Linux and Windows scheduled jobs synchronized.
% end extended example

% begin stoch-rx
\section{Stochastic Regular Expressions}
\label{sec:stoch-rx}
Important to our algorithm is finding the expected number of bits needed to
recover a string in one data source from a synchronized string in the other data
source. To do this, we must first develop a probability model for our language.
We do this with \emph{stochastic regular expressions}, regular expressions
annotated with probability information. With this, we can jointly express a
language and a probability distribution over that language.

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
% end stoch-rx

\section{Simple Symmetric Lenses}
\label{sec:ssl}
Classically, asymmetric lenses comprise three functions, \Put{}, \Get{}, and
\Create{}. These lenses only allow information to be lost on one side. We
enhance this style of lens to symmetric lenses with \emph{simple
  symmetric lenses}. Like the classical formulation of asymmetric lenses, simple
symmetric lenses have no complement. Simple symmetric lenses comprise 4
functions, \CreateR, \CreateL, \PutR, and \PutL. These functions satisfy the
following round-tripping laws:
\begin{equation}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x
\end{equation}
\begin{equation}
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y
\end{equation}
\begin{equation}
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{equation}
